Using algorithmic skeletons in EcmaScript to parallelize computation in browsers through Web Workers

### 2015-05-09 docs/paper/paper.tex
Some feedback from Rosita

% a short form should be given in case it is too long for the running head
\titlerunning{Algorithmic skeletons to parallelizing computation in browsers}

@@

Our library (p-j-s) only works with \verb+SharedTypedArray+s (and \verb+TypedArray+s) due to performance reasons discussed in Section~\ref{sec:overview}. It aims to provide a similar interface while not changing the native object's prototypes and dealing with the asynchronous nature of the operation. Given a \verb+SharedTypedArray+, the following code would do the same as the previous example but parallelizing the work:
\medskip

@@

To verify our hypothesis, we put together two benchmarks\footnote{All benchmarks in this document use the environment described in Tab.~\ref{tab:env}.} that aim to verify the difference between invoking \verb+postMessage+ with and without structured cloning for a \verb+SharedTypedArray+. Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one has 100000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one 1000000 (a million). As it can be seen from the  results in Tab.~\ref{tab:transf-vs-clone}, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects, it scales. On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth]{context}
\caption{Relative performance of parallel transformation compared to sequential one.}
\label{fig:context}
\end{figure}

@@

In this regard, the more complex the computation to be performed on each array element the better the parallel implementation will perform. That is because in the parallel case, the cost of this processing is dwarfed by the cost of the actual computation.

@@

\subsection{Results}
The results for the benchmark are the displayed in Tab.~\ref{tab:benchmark} and Figure~\ref{fig:relative} shows the relative performance of the parallel approach compared to the sequential one.
\begin{table}
  \centering
  \begin{tabular}{|r|c|c|}
    \hline
    Elements & Serial [ops/sec] & Parallel [ops/sec] \\
    \hline
    100 & \textbf{154442} & 377.2\\
    1000 & \textbf{17460.2} &  368.4 \\
    10000 & \textbf{1768.6} & 311.2 \\
    100000 & \textbf{179.2} & 172.2 \\
    1000000 & 16.96 & \textbf{33.2} \\
    \hline
  \end{tabular}
  \caption{Comparing p-j-s (parallel) and plain JS (serial) performance for an image transformation.}
  \label{tab:benchmark}
\end{table}

@@

During our research we found that browser memory usage greatly increases when using shared memory after transforming an image multiple times. Memory is relased after a particular time period (which seems to indicate there is not a leak), but while memory usage is high performance degrades. We plan to spend some time on this issue.\\

@@

>>>>

% a short form should be given in case it is too long for the running head
\titlerunning{Algorithmic skeletons to parallelize computation in browsers}

@@

Our library (p-j-s) only works with \verb+SharedTypedArray+s (and \verb+TypedArray+s) due to performance reasons discussed in Section~\ref{sec:implementation}. It aims to provide a similar interface while not changing the native object's prototypes and dealing with the asynchronous nature of the operation. Given a \verb+SharedTypedArray+, the following code would do the same as the previous example but parallelizing the work:
\medskip

@@

To verify our hypothesis, we put together two benchmarks\footnote{All benchmarks in this document use the environment described in Tab.~\ref{tab:env}.} that aim to verify the difference between invoking \verb+postMessage+ with and without structured cloning for a \verb+SharedTypedArray+. Both transfer a \verb+SharedTypedArray+ back and forth between the UI thread and a worker; the only difference between the two is that one has 100,000 (a hundred thousand) elements in the \verb+SharedTypedArray+ and the other one 1,000,000 (a million). As it can be seen from the  results in Tab.~\ref{tab:transf-vs-clone}, increasing the amount of elements by 10x does not change the amount of operations that can be performed when using \verb+Transferable+ objects, it scales. On the other hand, the amount of operations that can be performed with structured cloning greatly decreases.

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth]{context}
\caption{Example of global and local context being merged for chain steps.}
\label{fig:context}
\end{figure}

@@

In this regard, the more complex the computation to be performed on each array element the better the parallel implementation will perform. The reason for this is that, in the parallel case, the cost of this processing is dwarfed by the cost of the actual computation.

@@

\subsection{Results}
The results for the benchmark are the displayed in Tab.~\ref{tab:benchmark} and Figure~\ref{fig:relative} shows the relative performance of the parallel approach compared to the sequential one.
\begin{table}
  \centering
  \begin{tabular}{|r|c|c|}
    \hline
    Elements & Serial [ops/sec] & Parallel [ops/sec] \\
    \hline
    100 & \textbf{154442} & 377.2\\
    1,000 & \textbf{17460.2} &  368.4 \\
    10,000 & \textbf{1768.6} & 311.2 \\
    100,000 & \textbf{179.2} & 172.2 \\
    1,000,000 & 16.96 & \textbf{33.2} \\
    \hline
  \end{tabular}
  \caption{Comparing p-j-s (parallel) and plain JS (serial) performance for an image transformation.}
  \label{tab:benchmark}
\end{table}

@@

During our research we found that browser memory usage greatly increases when using shared memory after transforming an image multiple times. Memory is relased after a particular time period (which seems to indicate there is not a leak), but while memory usage is high, performance degrades. We plan to spend some time on this issue.\\

@@

\bibitem{promise} Promise, MDN, \url{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Reference/Global_Objects/Promise}
